---
title: "Assignment4"
author: "Sindhu Vegiraju"
date: "21/03/2021"
output: html_document
---
Read the Pharmaceuticals dataset to Pharma
```{r}
Pharma<-read.csv("D:/MACHINE LEARNING/Assignment4/Pharmaceuticals.csv")
head(Pharma)
```
```{r}
#a)Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.
```
Data Preparation

Prior to clustering data, remove or estimate missing data and rescale variables for comparability.
```{r}
x <- na.omit(Pharma)# listwise deletion of missing
```
collecting only the quantitative variables(1-9) to cluster the 21 firms
```{r}
Pharma1<-Pharma[,3:11]
head(Pharma1)
```
Scale all the quantitative variables in the dataframe (Z-score)
```{r}
Pharma2<-scale(Pharma1)
head(Pharma2)
```
Determining the number of clusters to do the cluster analysis

Calculating number of clusters using Elbow Method 

install.packages("factoextra")
```{r}
library(factoextra)   
fviz_nbclust(Pharma2, kmeans, method = "wss") + labs(subtitle = "Elbow Method")
```

Silhouette method for determining number of clusters

```{r}
fviz_nbclust(Pharma2, kmeans, method = "silhouette")+ labs(subtitle = "Silhouette Method")
```

With the plots shown above,it is clear that the number of clusters are 5 and they are enough to show the variations in the data

```{r}
set.seed(1)
k5<- kmeans(Pharma2,centers=5,nstart = 25) 
#Visualize the output
k5$centers #centroids
fviz_cluster(k5,data = Pharma2)#Visualize the clusters
```

K-Means Cluster Analysis- Fit the data with 5 clusters

```{r}
fit<-kmeans(Pharma2,5)
```

Finding the mean value of all quantitative variables for each cluster

```{r}
aggregate(Pharma2,by=list(fit$cluster),FUN=mean)
```
append cluster assignment
```{r}
Pharma3<-data.frame(Pharma2,fit$cluster)
Pharma3
```
Visualize the cluster plot
```{r}
library(cluster)
clusplot(Pharma2,fit$cluster,color = TRUE,shade = TRUE,labels = 2,lines = 0)
```
```{r}
#b)Interpret the clusters with respect to the numerical variables used in forming the clusters.
```
Cluster 1 - Row 8,9,12,14

Cluster 2 - Row 1,4,7,10,16,19,21

Cluster 3 - Row 2,6,18

Cluster 4 - Row 3,5,20

Cluster 5 - Row 11,13,15,17

By observing the mean values of all quantitative variables for each cluster

Cluster 1 has the lowest Market_Cap, highest Beta,lowest PE_Ratio,highest Leverage, highest Rev_Growth.

Cluster 2 has the lowest Rev_Growth, highest Net_Profit_Margin

Cluster 3 has the highest PE_Ratio,lowest ROE,lowest ROA,lowest Net_Profit_Margin

Cluster 4 has the lowest Beta,lowest Asset_Turnover

Cluster 5 has the highest Market_Cap, highest ROE, highest ROA,highest Asset_Turnover,lowest Leverage
```{r}
#c)Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)
```  
There is a pattern in the clusters with respect to Media recommendation variable.

Cluster 1 with highest Beta, highest Leverage, highest Rev_Growth has mostly Moderate Buy Recommendation.

Cluster 2 with highest Net_Profit_Margin has mostly Hold Recommendation

Cluster 3 with highest PE_Ratio has Hold Recommendation

Cluster 5 with highest Market_Cap, highest ROE, highest ROA, highest Asset_Turnover has equal Hold and Moderate Buy Recommendation

Cluster 4 with lowest Asset_Turnover has the Strong Buy Recommendation

Could see a pattern among the clusters with respect to variables(10 to 12)

Clusters 1,4 has mostly Moderate Buy Recommendation

Clusters 2,3 has mostly Hold Recommendation
```{r}                                                        
#d)Provide an appropriate name for each cluster using any or all of the variables in the dataset.
```
Cluster1 - low Market_Cap, high Beta,low PE_Ratio, high Leverage, high Rev_Growth cluster

Cluster2 - low Rev_Growth, high Net_Profit_Margin cluster

Cluster3 - high PE_Ratio,low ROE,low ROA,low Net_Profit_Margin cluster

Cluster4 - low Beta,low Asset_Turnover cluster

Cluster5 - high Market_Cap, high ROE, high ROA, high Asset_Turnover,low Leverage cluster





